Quasi-Monte Carlo Methods: What, Why, and How?

### 2025-01-08 HickernellMCQMC2024_QMCTutorial.tex
Updates from Overleaf

Since the low discrepancy sequences that drive qMC are defined on the unit cube, $[0,1]^d$, we perform a variable transformation $\bst = \bigl (\Phi^{-1}(x_1), \ldots, \Phi^{-1}(x_d) \bigr )/\sqrt{2}$, where is $\Phi$ is the cumulative distribution function of the standard Gaussian random variable.  This reimagines the integral $I$ as the expectation of a function, $f$, of a standard uniform random variable:
\begin{multline}\label{eq:keisterB}
	\mu = \int_{[0,1]^d} \underbrace{\pi^{d/2} \cos\Bigl (\sqrt{2} \bigl \lVert \bigl( \Phi^{-1}(x_1), \ldots, \Phi^{-1}(x_d)\bigr) \bigr\rVert \Bigr)}_{=:f(\bsx)} \, \rmd \bsx \\
	= \bbE[Y]
	= \bbE[f(\bsX)], \qquad \bsX \sim \calU[0,1]^d.
\end{multline}

@@

\begin{figure}
	\centering
	\includegraphics[width =10cm]{\figpath/n_is_16777216_d_is_6_n_rep_is_50_KeisterErrors.pdf}
	\caption{The relative error of approximationg $\mu$ defined in \eqref{eq:keisterB} by the sample mean, $\hat{\mu}_n$ defined in \eqref{eq:sample_mean} for various choices of nodes, $\bsX_0, \bsX_1, \ldots$.  Grids have the largest error and low discrepancy (qMC) nodes have the smallest error. \label{fig:keister-err}}
\end{figure}

@@

	\item \emph{Low discrepancy nodes.}  For qMC methods the error decays nearly like $\mathcal{O}(n^{-1})$, which for this example correponds to a reduction in error of several orders of magnitude for large enough $n$.  Typically, low discrepancy sequences have preferred sample sizes.  For this Sobol' sequence, $n$ an positive integer power of $2$ tends to give better accuracy than arbitrary $n$.

>>>>

Since the low discrepancy sequences that drive qMC are defined on the unit cube, $[0,1]^d$, we perform a variable transformation $\bst = \bigl (\Phi^{-1}(x_1), \ldots, \Phi^{-1}(x_d) \bigr )/\sqrt{2}$, where is $\Phi$ is the cumulative distribution function of the standard Gaussian random variable.  This reimagines the integral $I$ as the expectation of a function, $f$, of a standard uniform random variable:
\begin{multline}\label{eq:keisterC}
	\mu = \int_{[0,1]^d} \underbrace{\pi^{d/2} \cos\Bigl (\sqrt{2} \bigl \lVert \bigl( \Phi^{-1}(x_1), \ldots, \Phi^{-1}(x_d)\bigr) \bigr\rVert \Bigr)}_{=:f(\bsx)} \, \rmd \bsx \\
	= \bbE[Y]
	= \bbE[f(\bsX)], \qquad \bsX \sim \calU[0,1]^d.
\end{multline}

@@

\begin{figure}
	\centering
	\includegraphics[width =10cm]{\figpath/n_is_16777216_d_is_6_n_rep_is_50_KeisterErrors.pdf}
	\caption{The relative error of approximationg $\mu$ defined in \eqref{eq:keisterC} by the sample mean, $\hat{\mu}_n$ defined in \eqref{eq:sample_mean} for various choices of nodes, $\bsX_0, \bsX_1, \ldots$.  Grids have the largest error and low discrepancy (qMC) nodes have the smallest error. \label{fig:keister-err}}
\end{figure}

@@

	\item \emph{Low discrepancy nodes.}  For qMC methods the error decays nearly like $\mathcal{O}(n^{-1})$, which for this example correponds to a reduction in error of several orders of magnitude for large enough $n$.  Typically, low discrepancy sequences have preferred sample sizes.  For this Sobol' sequence, $n$ a positive integer power of $2$ tends to give better accuracy than arbitrary $n$.
